Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-08-02 paper.tex
3Aug refinement v2

\subsection{Stability Selection}
Stability selection is a subsampling strategy with a wide range of applications.  In the context of regularisation methods, it is a way of estimating, 
empirically, the significance of the effects. Compared to permutation, the amount of computation is significantly less.  
Analyses are based on only a fraction of the data  and the regularisation parameter need not be optimised.  
To perform stability selection, we did the following.  
For LMM-Lasso and glmnet, the regularisation parameter was fixed to a "suitable" value. This was done  
via a preliminary analysis of the data where we adjusted the regularisation parameter  to yield
between 20 to 30 SNPs with non-zero effect sizes.  Here, the number of non-zero effects is rather arbitrary and any reasonable 
number of non-zero effects would suffice. 
 We then repeatedly subsampled, without replacement, the data. As recommended \cite{meinshausen2010stability}, we drew 100 subsamples of size $n/2$ from the data. Each subsample was analysed with LMM-Lasso and glmnet, with their regularisation parameters fixed. 
A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
number of times the SNP had a non-zero effect size over all the replicates divided by the number of replicates (which was 100).

 To calculate, empirically,  the statistical significance of each 
SNP across the genome, we counted the number of times a SNP had a non-zero effect size over all the replicates. We then 
divided this number by the number of replicates (which was 100) to form an empirical probability estimate of the significance of a SNP.

====>

When using BigRR, LMM-Lasso, and glmnet, the results are affected by the amount of regularization ($\lambda$). 
BigRR has an approximate way of setting each SNP's separate regularization parameter (ref). However, the optimal value of 
$\lambda$ must be found when using LMM-Lasso and glmnet. Also, 
all three applications yield the effect sizes of the SNP across the entire genome but not their statistical significance. To address 
these issues, we made use of stability selection \cite{meinshausen2010stability}. 
Stability selection is a resampling strategy.  With stability selection, we were able to avoid having to optimise $\lambda$ while 
still being able to calculate, empirically, the statistical significance of the SNP effects.

We performed stability selection as follows. For LMM-Lasso and glmnet, we performed a preliminary analysis to find an 
appropriate value for the regularization parameter.  We adjusted $\lambda$ so that LMM-Lasso and glmnet yielded 
between 10 to 30 SNPs with non-zero effects. Fortunately, with stability selection, the setting of $\lambda$ does not have to be exact 
 We then repeatedly subsampled, 
without replacement, the data. As recommended \cite{meinshausen2010stability}, we draw 100 subsamples of size $n/2$. We 
analysed the subsample, with $\lambda$ fixed for LMM-Lasso and glmnet.  To calculate, empirically,  the statistical significance of each 
SNP across the genome, we counted the number of times a SNP had a non-zero effect size over all the replicates. We then 
divided this number by the number of replicates (which was 100) to form an empirical probability estimate of the significance of a SNP.

For BigRR, we modified our stability selection procedure slightly. 
There was no need to find an appropriate value for the regularization 
parameters as the BigRR package has an internal procedure for their 
calculation. We draw subsamples as above and analyzed the data with BigRR. However, within an analysis, we ordered 
the SNPs according to the absolute size of their SNP effects and recorded the top 20 SNPs. We then measured the significance 
of the SNPs across the entire genome as above.

>>>>

\subsection{Stability Selection}
Stability selection \cite{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of significance in regression models where 
the number of parameters ($p$) can be much larger than the number of samples ($n$). We chose stability selection over permutation and other sampling procedures because of its low computational cost. We employed stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.

To obtain significance estimates via stability selection, we did the following. For LMM-Lasso and glmnet, we began by performing a preliminary analysis of a replicate where  we tuned the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
This tuning of the regularisation parameter was done for data under each simulation scenario but it was not necessary to do it for every replicate within a scenario. It was sufficient to pick a single replicate at random upon which to tune the regularisation parameter. 
Here, we had the luxury of knowing the genetic conditions 
under which the simulated trait data were generated. We knew that 20 to 30 SNP-trait associations was a reasonable number of findings to expect. 
 However, stability selection estimates are robust  \cite{meinshausen2010stability} 
 and the regularisation parameter could have been tuned to any reasonable number of 
 non-zero effects. We sampled repeatedly, without replacement, from the replicate. We drew 100 replicate subsets of size $n/2$.  
Each replicate subset was then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
number of times the SNP had a non-zero effect size over all the replicates divided by the number of replicates (which was 100).

For BigRR, we modified our stability selection procedure slightly. 
There was no need to tune the regularisation parameter for BigRR as an optimal value is found as part of its analysis. 
As described above, we drew 100 replicate subsets of size $n/2$ and analysed these data with BigRR. We then ordered the 
SNPs according to the absolute size of their SNP effects and recorded the top 20 SNPs. 
A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
number of times the SNP was in the top 20 SNPs divided by the number of replicates.
